\chapter{Exam}
\label{chap:Exam}

The question in all courses is: ``What is relevant for the exam?''

Each semester I will try to give an answer to that question again and again \ldots and again. \textbf{Here is my answer (not fixed yet) for the winter term 2016/17}:

You should be able to give the right answer to the following questions and be able to fullfil the given tasks:

\section{Questions: Foundations}
\label{sec:questions_foundations}

\begin{itemize}
 \item What is an agent?
 \item What is an autonomous agent?
 \item Sketch a state-of-the-art robot control architecture and explain its components.
 \item What is the difference between trueness and precision?
 \item Name two different sensors and explain their categories, properties, and value properties.
 \item What is an actuator?
 \item What is odometry?
\end{itemize}

\section{Questions: External Software and Developing Tools}
\label{sec:questions_tools}

\begin{itemize}
 \item What is a version control system and what is it good for?
 \item What does git pull, push, commit, etc. do?
 \item What is ROS and what is it good for?
 \item Describe the publish-subscribe model for communication.
 \item Describe the notion of nodes, topics, messages, publishers, subscribers.
 \item What is a Build Chain?
 \item What is CMake and for what do we use it?
\end{itemize}

\section{Questions: C++}
\label{sec:questions_c++}

\begin{itemize}
 \item What is written in the header file and what in the source file of a C++ class?
 \item How does a shared\_ptr work and what is it good for?
 \item Explain roughly what the gcc compiler does.
\end{itemize}

\section{Questions: Real Robots}
\label{sec:questions_robots}

\begin{itemize}
 \item 
\end{itemize}

\section{Questions: Gazebo Simulation}
\label{sec:questions_simulation}

\begin{itemize}
 \item Why do we use the logical camera sensor in simulation?
 \item Denote the input and the output of AMCL and sketch how AMCL is calculating the robtos position based on the input.
\end{itemize}
