\chapter{Introduction}
\label{chap:Introduction}

\section{Purpose of this Document}
\label{sec:Purpose}

Instead of giving the students a lot of Powerpoint slides that compress the complex topics of this course into a bunch of bullet points, we wanted the students to have a dedicated document for preparing their exam. Unfortunately, such a document is a lot more work and it needs several iterations of thorough reviewing. Therefore, we came up with the compromise that the students write this document theirselves and the lecturers only have to review and improve it. 

Maybe you ask: "What is wrong with Powerpoint?" 

A Powerpoint presentation should aid a presenter in presenting a certain message or information to his audience. The slides alone often miss critical parts (that couldn't easily condensed into bullet points) and are unclear without the corresponding vocal line of the presenter. Nevertheless, students end up alone with the slides trying to prepare their examinations.

Apart from that, try to google topics like "disadvantages of powerpoint" or "why we should ban Powerpoint" and you will find interesting articles that connect Powerpoint and the Accident of the Columbia Space Shuttle [\href{http://www.nytimes.com/2003/12/14/magazine/2003-the-3rd-annual-year-in-ideas-powerpoint-makes-you-dumb.html}{1}] [\href{http://www.edwardtufte.com/bboard/q-and-a-fetch-msg?msg_id=0001yB&topic_id=1&topic=Ask+E%2eT%2e}{2}].

\section{How to Contribute to this Document?}
\label{sec:Contribute}

Basically there are two aspects. At first you need to get the \LaTeX\  source code of this document from our \href{https://github.com/carpe-noctem-cassel/cnc-turtlebots}{cnc-turtlebots GitHub Repository}. It is located in the folder \verb#doc/latex#. Anybody can download it, but for making changes to it you need to have an GitHub-Account which was given the corresponding previliges. That is typically arranged during the first session of this course. 

This leaves us with the requirements, that you are able to write \LaTeX\ and know how to use GIT in combination with GitHub. Regarding GIT and GitHub we recommend to read Section~\ref{sec:Git}. 

Maybe \LaTeX\ is hard to master when it comes to tables, graphics and customizing the layout of a document, but you don't have to do that for contributing to this document. You should only be able to able to write text referencing images and place useful links from time to time. Furthermore, you should be able to structure the content into chapters, sections, and subsections. Everything just mentioned is already done in this document, so you can start by copying commands from the corresponding sections. 

For further reference and an in-depth study of \LaTeX, we recommend the following sources:

\begin{description}
  \item[\href{http://www.wiwiss.fu-berlin.de/fachbereich/vwl/iso/links/latex_einfuehrung_manuela_juergens.pdf}{\LaTeX\ \textendash\ eine Einf\"uhrung und ein bisschen mehr}:] A little bit longer introduction to \LaTeX\ basics.
  \item[\href{https://en.wikibooks.org/wiki/LaTeX}{\LaTeX\ WikiBook}:] Nice for looking up simple stuff. 
  \item[\href{https://tex.stackexchange.com}{tex.stackexchange.com}:] Stack Overflow for \LaTeX
  \item[\href{http://ftp.fau.de/ctan/graphics/pgf/base/doc/pgfmanual.pdf}{TikZ}:] \LaTeX package for designing advanced graphics directly in \LaTeX.
\end{description}

The final issue is to decide which editor you choose for compiling this \LaTeX document. We recommend \href{http://www.xm1math.net/texmaker/}{TexMaker}, because it is available for Ubuntu and Windows, although, there currently exists an issue with short cuts under Ubuntu 16.04 LTS. Consider this \href{http://askubuntu.com/questions/786280/texmaker-shortcuts-not-working-on-ubuntu-16-04}{Ask Ubuntu Post} for solving it.

With TexMaker you only have to open the \verb#Software_Reference_Book.tex# file and click on \verb#Quick Build#. If you get compile errors, you probably need to install the necessary \LaTeX packages. Under Ubuntu this is done by executing the following console command:

\verb#sudo apt-get install texlive-full#
